---
title: "SES 5215 | Analytic Methods of Urban Planning: Quantitative"
author: "Sai Joshi, Jialei Tang, Yuanzhao Wang "
date: "2021/10/22"
output: 
  rmdformats::downcute
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```

```{r, warning = FALSE, message = FALSE, echo = TRUE, results = TRUE}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
library(tigris)
library(sf)
library(gridExtra)
library(jtools)
library(interactions)
```
# Assignment 1 & 2
# Research Question

Does the environnment sustainability status of a city impact the wellbeing of its citizens?

# Citation articles

C. Yin & C. Shao. Revisiting commuting, built environment and happiness: New evidence on a nonlinear relationship,
*Transportation Research Part D: Transport and Environment* Volume 100 (2021) https://doi.org/10.1016/j.trd.2021.103043.

Riffat, S., Powell, R. & Aydin, D. Future cities and environmental sustainability. *Fut Cit & Env 2, 1* (2016). https://doi.org/10.1186/s40984-016-0014-2

# Sample population

The cities of which data on wellbeing, emissions, energy mix, vulnerability assessment and  could be obtained.

# Variables

## Categorical Variables (2)

* Vulnerability Assessment 
* C40 Status

## Continuous Variables (4)

* Emissions
* Wellbeing index 
* Population density
* Percentage of Energy Use from Renewable Sources

# Number of observations (rows) in our data set

```{r, warning = FALSE, message = FALSE, echo = TRUE, results = TRUE}
main <- st_read("values_variablesonly.xlsx")
total <- st_read("combinedmain_valuesonlyr.csv")
```

Number of Observations: 

291 Observations (Rows/Cities)

In our preliminary round of research, we identified 45 out of 292 cities that can provide us with data on the well-being index. There are 9 cities which did not have total carbon emissions reported. We accounted for the gaps by gathering the emissions numbers for the cities that did not have that data in the first data set. 

We intend to collate further data to include in the event that the number of cities are insufficient for analysis.
```{r, warning = FALSE, message = FALSE, echo = TRUE}
kable(head(total))
```
```{r, warning = FALSE, message = FALSE, echo = TRUE}
kable(head(main))
```
# Reading our data

## Continuous variables:

* Emissions 
* Well-being index 
* Population density
* Energy Mix (Source of electricity) 

```{r read xlsx, warning = FALSE, message = FALSE, echo = TRUE, results = TRUE}
data <- read_xlsx(path = "values_variablesonly.xlsx", 
                      sheet = "Sheet1") %>%
  mutate(log_emissions = log(TotalEmissions))
energy_t_test <- t.test(data$PercentRenew)
dens_t_test <- t.test(data$PopulationDensity)
emission_t_test <- t.test(data$TotalEmissions)
wellbeing_t_test <- t.test(data$WellBeing)
```                  


```{r calculate continuous variables, warning = FALSE, message = FALSE, echo = TRUE, results = TRUE}

energy_quartiles <- quantile(data$PercentRenew, na.rm = TRUE)
dens_quartiles <- quantile(data$PopulationDensity, na.rm = TRUE)
emission_quartiles <- quantile(data$TotalEmissions, na.rm = TRUE)
wellbeing_quartiles <- quantile(data$WellBeing, na.rm = TRUE)

energy_dev <- sd(data$PercentRenew, na.rm = TRUE)
dens_dev <- sd(data$PopulationDensity, na.rm = TRUE)
emission_dev <- sd(data$TotalEmissions, na.rm = TRUE)
wellbeing_dev <- sd(data$WellBeing, na.rm = TRUE)

energy_hist <- ggplot(data) +
  geom_histogram(aes(x = PercentRenew),
                 bins = 30)

dens_hist <- ggplot(data) +
  geom_histogram(aes(x = PopulationDensity),
                 bins = 30) +
  scale_x_continuous(trans = "log")

emission_hist <- ggplot(data) +
  geom_histogram(aes(x = log_emissions),
                 bins = 30)

wellbeing_hist <- ggplot(data) +
  geom_histogram(aes(x = WellBeing),
                 bins = 30)
```


```{r display continuous variables, warning = FALSE, message = FALSE, echo = TRUE, results = TRUE}

cont_summary <- tibble(`Variable` = c("Percentage of Energy Use from Renewable Sources", 
               "Population density (people per square kilometers)", 
               "Total Emissions (metric tonnes CO2e)",
               "Well-being scores"),

`Sample mean` = c(energy_t_test$estimate,
                    dens_t_test$estimate,
                    emission_t_test$estimate,
                    wellbeing_t_test$estimate* 1000),

`Population mean (95% confidence) - low` = 
    c(energy_t_test$conf.int[1],
      dens_t_test$conf.int[1],
      emission_t_test$conf.int[1],
      wellbeing_t_test$conf.int[1]* 1000),
`Population mean (95% confidence) - high` =
    c(energy_t_test$conf.int[2],
      dens_t_test$conf.int[2],
      emission_t_test$conf.int[2],
      wellbeing_t_test$conf.int[2] * 1000),
`Median` = c(energy_quartiles[3],
             dens_quartiles[3],
             emission_quartiles[3],
             wellbeing_quartiles[3] * 1000),
`Interquartile range` = c(energy_quartiles[4] - energy_quartiles[2],
                            dens_quartiles[4] - dens_quartiles[2],
                            emission_quartiles[4] - emission_quartiles[2],
                            wellbeing_quartiles[4] - wellbeing_quartiles[2]* 1000),
`Standard deviation` = c(energy_dev,
                          dens_dev,
                          emission_dev,
                          wellbeing_dev* 1000))

kable(cont_summary, digits = 0)
```

```{r display histograms, warning = FALSE, message = FALSE, echo = TRUE, results = TRUE}

pretty_energy_hist <- energy_hist +
  theme_bw() +
  scale_x_continuous(name = "Percentage of Energy Use\nfrom Renewable Sources") +
  scale_y_continuous(name = "Number of cities") +
  theme(axis.text.x = element_text(angle = 90))

pretty_dens_hist <- dens_hist +
  theme_bw() + 
  scale_x_continuous(name = "Population density\n(residents per square kilometer)",
                     trans = "log",
                     breaks = c(0.1, 1, 10, 100, 1000, 10000),
                     labels = c("0.1", "1", "10", "100", "1000", "10,000")) +
  scale_y_continuous(name = "Number of cities") +
  theme(axis.text.x = element_text(angle = 90))

pretty_emission_hist = emission_hist +
  theme_bw() +
  scale_x_continuous(name = "Total Emissions (log(Metric Tonnes CO2e))") +
  scale_y_continuous(name = "Number of cities") +
  theme(axis.text.x = element_text(angle = 90)) 

pretty_wellbeing_hist = wellbeing_hist +
  theme_bw() +
  scale_x_continuous(name = "Well-being scores") +
  scale_y_continuous(name = "Number of cities") +
  theme(axis.text.x = element_text(angle = 90))

grid.arrange(pretty_energy_hist, pretty_dens_hist, pretty_emission_hist, pretty_wellbeing_hist,
             ncol = 2, nrow = 2)
```

# Sample Proportion for each Category

## Categorical variables:

* Vulnerability Assessment
* C40 Status

```{r}
pct_C40 <- t.test(main$C40_Status == "C40")
pct_nonC40 <- t.test(main$C40_Status == "No")
```

```{r}
cat_summary_C40 <- tibble(`C40 Cities` = 
                              c("C40",
                                "Non C40"),
                            `Sample proportion` = 
                              c(pct_C40$estimate * 100,
                                pct_nonC40$estimate *100),
                            `95-percent confidence - low` = 
                              c(pct_C40$conf.int[1] * 100,
                                pct_nonC40$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_C40$conf.int[2] * 100,
                                pct_nonC40$conf.int[2] * 100))

kable(cat_summary_C40, digits = 0)
```

# The 95-percent confidence interval for the population proportion in each category

```{r}
pct_VAdone <- t.test(main$Vadone == "Yes")
pct_VAinprogress <- t.test(main$Vadone == "In progress")
pct_VAno <-  t.test(main$Vadone == "No")
```

```{r}

cat_summary_VAdone <- tibble(`Conducted Vulnerability Assessment?` = 
                              c("Yes",
                                "In Progress",
                                "No"),
                            `Sample proportion` = 
                              c(pct_VAdone$estimate * 100,
                                pct_VAinprogress$estimate * 100,
                                pct_VAno$estimate * 100),
                            `95-percent confidence - low` = 
                              c(pct_VAdone$estimate * 100,
                                pct_VAinprogress$estimate * 100,
                                pct_VAno$estimate * 100),
                            `95-percent confidence - high` = 
                              c(pct_VAdone$conf.int[2] * 100,
                                pct_VAinprogress$conf.int[2] * 100,
                                pct_VAno$conf.int[2] * 100))

kable(cat_summary_VAdone, digits = 0)
```

# Assignment 3
# Bivariate analysis
## Correlations / regression on continuous variables
Well-being is our key variable. The components of the well-being index are

* Annual hours of sunshine hours
* Access to green space
* Crime
* Traffic congestion
* Happiness
* Quality of healthcare
* Work-life Balance
* Governance 

We recognize the limitations of this index and remain critical of its representation of the quality of life of citizens in the cities examined. Nevertheless, it is an interesting study and we are curious if the continuous and categorical variables we've identified can give the index more context as we establish its relationships to these variables.

In this exercise, we would like to investigate the correlations between 

* well-being and emissions
* well-being and population density
* well-being and energy-mix
* population and emissions 
* emissions and energy-mix

## Well-being and emissions

The p-value is more than 0.05 at 0.3891, which means the relationship is insignificant. This is also clear from the fact that the entire 95-percent confidence interval is both positive and negative, which suggests that the relationship is neither inverse or direct.
```{r}
cor.test(~ WellBeing + TotalEmissions, data = main)
```

```{r}
wb_emissions_model <- lm(WellBeing ~ TotalEmissions, data = main)

summary(wb_emissions_model)
```

```{r, warning = FALSE, message = FALSE, echo = FALSE, results = TRUE}
ggplot(main) +
  geom_point(aes(x = TotalEmissions, 
                 y = WellBeing,
                 size = TotalEmissions),
             alpha = 0.25) +
  labs(y="Well-being Score", x = "Total Emissions (Metric Tonnes CO2e)", size = "Total Emissions\n(Metric Tonnes CO2e)") +
  theme_bw()
```


## Well-being and population density

The p-value is more than 0.05 at 0.3906, which means the relationship is insignificant. This is also clear from the fact that the entire 95-percent confidence interval is both positive and negative, which suggests that the relationship is neither inverse or direct.
```{r}
cor.test(~ WellBeing + PopulationDensity, data = main)
```

```{r}
wb_popdens_model <- lm(WellBeing ~ PopulationDensity, data = main)

summary(wb_popdens_model)
```

## Well-being and energy mix

The p-value is less than 0.05 at 0.000009851, which means the relationship is significant. This is also clear from the fact that the entire 95-percent confidence interval is positive, which suggests that the relationship is direct.

This result is interesting and it could be suggesting that cleaner cities provide better conditions for a higher wellbeing.

```{r}
cor.test(~ WellBeing + PercentRenew, data = main)
```

```{r}
wb_energy_model <- lm(WellBeing ~ PercentRenew, data = main)

summary(wb_energy_model)
```


```{r, warning = FALSE, message = FALSE, echo = FALSE, results = TRUE}
ggplot(main) +
  geom_point(aes(x = PercentRenew, 
                 y = WellBeing),
             alpha = 0.25) +
  labs(y="Well-being Score", x = "Percentage of Energy Use from Renewable Sources") +
  theme_bw()
```

## Population density and emissions

The p-value is more than 0.05 at 0.4849, which means the relationship is insignificant. This is also clear from the fact that the entire 95-percent confidence interval is both positive and negative, which suggests that the relationship is neither inverse or direct.
```{r}
cor.test(~ PopulationDensity + TotalEmissions, data = main)
```

```{r}
popdens_te_model <- lm(PopulationDensity ~ TotalEmissions, data = main)

summary(popdens_te_model)
```

## Emissions and energy mix

The p-value is less than 0.05 at 0.0105, which means the relationship is significant. This is also clear from the fact that the entire 95-percent confidence interval is negative (it does not include zero), which suggests that the relationship is inverse.

```{r}
cor.test(~ TotalEmissions + PercentageRenewable, data = data)
```

```{r}
te_energy_model <- lm(TotalEmissions ~ PercentageRenewable, data = main)

summary(te_energy_model)
```


# Difference in means / regression on categorical variables

## Well-being and c40 status

The p-value is more than 0.05 at 0.2037, which means the relationship is insignificant. This is also clear from the fact that the entire 95-percent confidence interval is both positive and negative, which suggests that the relationship is neither inverse or direct.

```{r}
t.test(WellBeing ~ C40_Status, main)
```


```{r}
wb_c40_model <- lm(WellBeing ~ C40_Status, data = main)

summary(wb_c40_model)

```


## Well-being and status of vulnerability assessment

The p-value is less than 0.05 at 0.03 for the regression test, which means the relationship is significant. 
In order to run the t-test, a new column defined by whether Vulnerability Assessment was commenced was created, in which all "In progress" cities were changed to "Yes". Here, the p-value is 0.05957, which is greater than 0.05. It suggest that the relationship is not that significant between whether a city has commenced vulnerability assessment and its wellbeing. The 95 percent confidence interval also includes positive and negative values, which cannot point to the nature of the relationship (direct or inverse). 
This turns us to focus on the regression test of wellbeing and status of vulnerability assessment of cities (Yes, No, In progress), which is informs us that completion of the vulnerability assessment has a more significant relationship to wellbeing.

In this light, it appears that cities that have conducted vulnerability assessments are also cities whose citizens have a higher well being score. This could be due to the cities/governments which conduct vulnerability assessments are also more proactive in ensuring the well-being of their citizens.

```{r}
t.test(WellBeing ~ Vacommenced, data = main)
```


```{r}
wb_va_model <- lm(WellBeing ~ Vadone, data = main)

summary(wb_va_model)

```

# Assignment 4
# Multivariate Regression

```{r}
full_model <- lm(WellBeing ~ C40_Status + Vadone + PercentRenew + PopulationDensity + TotalEmissions, main)

summary(full_model)
```

In our bivariate regression, we derived the equation W = 5.932512 + 0.014506(PR), where W is the well-being score and PR is the percentage of energy use from renewable sources. This means that for percentage of energy use from renewable sources was associated with an increase of about 0.0014506 in the well-being score. When we control for c40 status, vulnerability assessment status, population density and total emissions, we find that relationship is weaker, although still significant with a p-value of 0.00011 compared to 0.000009851 in the bivariate regression. The relation is still positive, coherent with the 95% confidence interval. Ceteris paribus, each 1 percent increase in the percentage of energy use from renewable sources is associated with about 1.450e-02 increase in the well-being score. 

On the other hand, our bivariate regression (linear model) on well-being scores and completion of vulnerability assessment showed that the p-value is 0.03, which suggests that the two have a significant relationship.
Once we’ve controlled for our other variables, we find that relationship is actually weaker, in which the p-value is 0.26269 for cities that have completed vulnerability assessments. In this model, completion of vulnerability assessment appears to be associated to a decrease in well-being scores of -2.819e-01 (which is approximately -0.80529) but this relationship is insignificant given its high p-value.


The only significant relationship observed in the multivariate model is between the well-being score and the percentage of energy use from renewable sources. Apart from the well-being scores and completion of vulnerability assessment, relationships between well-being scores and other independent variables, namely C40 Status, population density and total emissions, cohere with our results from the earlier bivariate regression as insignificant.

# Assignment 5
# Transformations

## Linear Transformations

### Mean Centering 

In the earlier multivariate regression, the intercept predicted the well-being score for a C40 city while assuming zero values for emissions, population density, percentage of energy use from renewable sources, and a null value for vulnerability assessment status. This value is unrealistic. Through mean centering, we are able to predict a well-being score for a C40 city that assumes the average value of the other independent variables. 

A city that uses  less-than-average percentage energy from renewable sources would have a negative value and a city that uses more-than-average percentage energy from renewable sources would have a positive value. The same holds for population density.

The regression coefficents remain the same, but the intercept increased from 6.10 to 6.75. It refers to a C40 city average in terms of average percentage energy from renewable sources, population density, and total emissions. 

The R-Value does not change and remains at 0.45 in this linear transformation.
```{r}
centered_data <- main %>%
  mutate(pop_dens_centered = PopulationDensity - mean(PopulationDensity, na.rm=TRUE),
         TotalEmissions_centered = TotalEmissions - mean(TotalEmissions, na.rm=TRUE),
         PercentRenew_centered = PercentRenew - mean(PercentRenew, na.rm=TRUE))


centered_model <- lm(WellBeing ~ C40_Status + Vadone + pop_dens_centered + TotalEmissions_centered + PercentRenew_centered, centered_data)

coeff_names <- c("Constant" = "(Intercept)",
                 "Non-C40 City" = "C40_StatusNo",
                 "No vulnerability assessment done" = "VadoneNo",
                 "Conducted vulnerability assessment " = "VadoneYes",
                 "Percentage of Energy Use from Renewable Sources %" = "PercentRenew",
                 "Population Density (per sq km)" = "PopulationDensity", 
                 "Total Emissions  (metric tonnes CO2e)" = "TotalEmissions",
                 "Population Density (per sq km) Centered" = "pop_dens_centered",
                 "Total Emissions (metric tonnes CO2e) centered" = "TotalEmissions_centered",
                 "Percentage of Energy Use from Renewable Sources % Centered" = "PercentRenew_centered")

export_summs(full_model, centered_model, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial", "Centered"),
             coefs = coeff_names)
```

## Non-Linear transformations

### Logarithmic Transformation

In the preliminary examination of our dataset (see histograms in "Reading our data" section), we conducted a logarithmic transform for total emissions to visualize it in the histogram to see a better distribution.  We did not use the logged values in subsequent analysis. Here, we apply a logarithmic transformation in the multivariate regression. We used a base of 2 for the log, the interpretation of that coefficient will be the effect of doubling total emissions. 

```{r}
data_log <- centered_data %>%
  mutate(log_Emission = log(TotalEmissions), base = 2)

                   
centered_model_log <- lm(WellBeing ~ C40_Status + Vadone + PercentRenew_centered + pop_dens_centered + log_Emission, data_log)

coeff_names <- c("Constant" = "(Intercept)",
                 "Non-C40 City" = "C40_StatusNo",
                 "No vulnerability assessment done" = "VadoneNo",
                 "Conducted vulnerability assessment " = "VadoneYes",
                 "Percentage of Energy Use\nfrom Renewable Sources %" = "PercentRenew",
                 "Population Density (per sq km)" = "PopulationDensity", 
                 "Total Emissions\n(metric tonnes CO2e)" = "TotalEmissions",
                 "Population Density\n(per sq km) Centered" = "pop_dens_centered",
                 "Total Emissions\n(metric tonnes CO2e) centered" = "TotalEmissions_centered",
                 "Percentage of Energy Use\nfrom Renewable Sources % Centered" = "PercentRenew_centered",
                 "Total Emissions\n(metric tonnes CO2e) logged" = "log_Emission")

export_summs(full_model, centered_model, centered_model_log,
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial", 
                             "Centered",
                             "logged\nCentered"),
             coefs = coeff_names)
```

The R-value remains the same as the initial and centered models. Additionally, it is rather high at 0.45. This suggests that the model was quite good to begin with. The p value for total emissions improved slightly from 0.29 to 0.23, however, this still proves that there is no significant relationship between total emissions and well-being score within a 90 or 95 percent confidence interval. 

For practical purposes, we think that the mean centering model is more useful since it provides a practical prediction of a well-being score for a C40 city. On the other hand, The logarithmic transformation did not return a better model fit, and we would still not use the predicted well-being score based on total emissions as the relationship between the two is insignificant. 

```{r}
effect_plot(centered_model, pred = PercentRenew_centered, interval = TRUE) +
  scale_y_continuous(name = "Well-being Score") +
  scale_x_continuous(name = "Percentage of Energy Use from Renewable Sources (Centered)")
```

# Interaction

## Vulnerability Assessment and Total Emissions 

Based on the results above, the only significant relationship is between well-being scores and percentage of energy use from renewable sources. We are interested in whether the 1) completion of vulnerability assessment by a city and 2) C40 status will change that relationship.For this, we will use the logged model since we are investigating the percentage of energy from renewable sources. 

```{r}
coeff_names_interaction <- c("Constant" = "(Intercept)",
                 "Non-C40 City" = "C40_StatusNo",
                 "No vulnerability assessment done" = "VadoneNo",
                 "Conducted vulnerability assessment " = "VadoneYes",
                 "Population Density\n(per sq km) Centered" = "pop_dens_centered",
                 "Percentage of Energy Use\nfrom Renewable Sources % Centered" = "PercentRenew_centered",
                 "Total Emissions\n(metric tonnes CO2e) logged" = "log_Emission",
                 "Effect of non-completed VA" = "VadoneNo:PercentRenew_centered",
                 "Effect of completed VA" = "VadoneYes:PercentRenew_centered",
                 "Effect of non-C40 Status" = "C40_StatusNo:PercentRenew_centered")

interaction_model <- lm(WellBeing ~ C40_Status + Vadone + PercentRenew_centered + pop_dens_centered + log_Emission + Vadone*PercentRenew_centered, data_log)

interaction_model_2 <- lm(WellBeing ~ C40_Status + Vadone + PercentRenew_centered + pop_dens_centered + log_Emission + C40_Status*PercentRenew_centered, data_log)

export_summs(centered_model_log, interaction_model, interaction_model_2,
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("logged\nCentered",
                             "Effect of Completion of\nVulnerability Assessment",
                             "Effect of\nC40 Status"),
             coefs = coeff_names_interaction)
```


While the model  gives us a better fit when we interact the completion of vulnerability assessment with the percentage of energy use from renewable sources, the relationship is not strong given the high p values of 0.20 and 0.34. In other words, regardless of a city being in the progress of completing, has completed, or has not conducted vulnerability assessment, does not have a significant impact on the relationship between percentage of energy use from renewable sources and the well-being score. This insignificant relationship is shown in the graph below. 

```{r}
interact_plot(interaction_model, pred = PercentRenew_centered, modx = Vadone,
              interval = TRUE) +
  scale_x_continuous(breaks = breaks <- c(-40, -20, 0, 20, 40, 60, 80, 100),
                     labels = round(breaks * 1 + mean(data_log$PercentRenew_centered, na.rm = TRUE)),
                     name = "Percentage of Energy Use from Renewable Sources") +
  scale_y_continuous(name = "Predicted Well-being score")
```

The R2 value of the C40-percentage of energy use from renewable sources interaction model does not change from the original logged Centered model remaining at 0.45. The relationship is stronger and positive at 90 percent confidence level in C40 cities than in non-C40 cities, which has a negative and very close-to-zero coefficient. This relationship is visualized below. Interestingly, non-C40 cities seem to have a higher predicted well-being score than C40 cities irrespective of percentage of energy use from renewable sources.  

```{r}
interact_plot(interaction_model_2, pred = PercentRenew_centered, modx = C40_Status,
              interval = TRUE) +
  scale_x_continuous(breaks = breaks <- c(-40, -20, 0, 20, 40, 60, 80, 100),
                     labels = round(breaks * 1 + mean(data_log$PercentRenew_centered, na.rm = TRUE)),
                     name = "Percentage of Energy Use from Renewable Sources") +
  scale_y_continuous(name = "Predicted Well-being score")
```

