---
title: "SES 5215: Assignment 1"
author: "Sai Joshi, Jialei Tang, Yuanzhao Wang "
date: "2021/10/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```

```{r, warning = FALSE, message = FALSE, echo = TRUE, results = FALSE}
library(sf)
library(knitr)
```

# Research Question

Does the environnment sustainability status of a city impact the wellbeing of its citizens?

# Citation articles

C. Yin & C. Shao. Revisiting commuting, built environment and happiness: New evidence on a nonlinear relationship,
*Transportation Research Part D: Transport and Environment* Volume 100 (2021) https://doi.org/10.1016/j.trd.2021.103043.

Riffat, S., Powell, R. & Aydin, D. Future cities and environmental sustainability. *Fut Cit & Env 2, 1* (2016). https://doi.org/10.1186/s40984-016-0014-2

# Sample population

The cities of which data on wellbeing, emissions, energy mix, vulnerability assessment and water risks could be obtained.

# Variables

## Categorical Variables (2)

* Energy Mix (Source of electricity) ( > 30% renewable energy)
* Vulnerability Assessment (2017)
* Water Risks

## Continuous Variables (3)

* Emissions (2020)
* Wellbeing index (2020)
* Population density


# Number of observations (rows) in our data set

```{r, warning = FALSE, message = FALSE, echo = TRUE, results = FALSE}
main <- st_read("combined_r.csv")
total <- st_read("combinedmain_valuesonlyr.csv")
```

Number of Observations: 

291 Observations (Rows/Cities)

In our preliminary round of research, we identified 48 out of 292 cities that can provide us with data on the well-being index. There are some gaps in data for the other variables (namely emissions and water risks, which would bring the number of cities to 28) 

We intend to collate further data to include in the event that the number of cities are insufficient for analysis.
```{r, warning = FALSE, message = FALSE, echo = TRUE}
kable(head(total))
```
```{r, warning = FALSE, message = FALSE, echo = TRUE}
kable(head(main))
```




