---
title: "SES 5215: Assignment 1&2"
author: "Sai Joshi, Jialei Tang, Yuanzhao Wang "
date: "2021/10/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```

```{r, warning = FALSE, message = FALSE, echo = TRUE, results = FALSE}
library(tidyverse)
library(sf)
library(knitr)
library(readxl)
library(gridExtra)
```

# Research Question

Does the environnment sustainability status of a city impact the wellbeing of its citizens?

# Citation articles

C. Yin & C. Shao. Revisiting commuting, built environment and happiness: New evidence on a nonlinear relationship,
*Transportation Research Part D: Transport and Environment* Volume 100 (2021) https://doi.org/10.1016/j.trd.2021.103043.

Riffat, S., Powell, R. & Aydin, D. Future cities and environmental sustainability. *Fut Cit & Env 2, 1* (2016). https://doi.org/10.1186/s40984-016-0014-2

# Sample population

The cities of which data on wellbeing, emissions, energy mix, vulnerability assessment and  could be obtained.

# Variables

## Categorical Variables (2)

* Vulnerability Assessment 
* C40 Status

## Continuous Variables (4)

* Emissions
* Wellbeing index 
* Population density
* Energy Mix (Source of electricity) 

# Number of observations (rows) in our data set

```{r, warning = FALSE, message = FALSE, echo = TRUE, results = FALSE}
main <- st_read("values_variablesonly.xlsx")
total <- st_read("combinedmain_valuesonlyr.csv")
```

Number of Observations: 

291 Observations (Rows/Cities)

In our preliminary round of research, we identified 45 out of 292 cities that can provide us with data on the well-being index. There are some gaps in data for the other variables (namely emissions and water risks, which would bring the number of cities to 28) 

We intend to collate further data to include in the event that the number of cities are insufficient for analysis.
```{r, warning = FALSE, message = FALSE, echo = TRUE}
kable(head(total))
```
```{r, warning = FALSE, message = FALSE, echo = TRUE}
kable(head(main))
```
# Reading our data

## Our continuous variables are:

* Emissions 
* Wellbeing index 
* Population density
* Energy Mix (Source of electricity) 

```{r read xlsx, warning = FALSE, message = FALSE, echo = TRUE, results = FALSE}
data <- read_xlsx(path = "values_variablesonly.xlsx", 
                      sheet = "Sheet1")
energy_t_test <- t.test(data$PercentageRenewable)
dens_t_test <- t.test(data$PopulationDensity)
emission_t_test <- t.test(data$TotalEmissions)
wellbeing_t_test <- t.test(data$WellBeing)
```                  


```{r calculate continuous variables, warning = FALSE, message = FALSE, echo = TRUE, results = FALSE}

energy_quartiles <- quantile(data$PercentageRenewable, na.rm = TRUE)
dens_quartiles <- quantile(data$PopulationDensity, na.rm = TRUE)
emission_quartiles <- quantile(data$TotalEmissions, na.rm = TRUE)
wellbeing_quartiles <- quantile(data$WellBeing, na.rm = TRUE)

energy_dev <- sd(data$PercentageRenewable, na.rm = TRUE)
dens_dev <- sd(data$PopulationDensity, na.rm = TRUE)
emission_dev <- sd(data$TotalEmissions, na.rm = TRUE)
wellbeing_dev <- sd(data$WellBeing, na.rm = TRUE)

energy_hist <- ggplot(data) +
  geom_histogram(aes(x = PercentageRenewable),
                 bins = 30)

dens_hist <- ggplot(data) +
  geom_histogram(aes(x = PopulationDensity),
                 bins = 30) +
  scale_x_continuous(trans = "log")

emission_hist <- ggplot(data) +
  geom_histogram(aes(x = TotalEmissions),
                 bins = 30)

wellbeing_hist <- ggplot(data) +
  geom_histogram(aes(x = WellBeing),
                 bins = 30)
```


```{r display continuous variables, warning = FALSE, message = FALSE, echo = TRUE, results = FALSE}

cont_summary <- tibble(`Variable` = c("Energy Mix", 
               "Population density (people per square kilometers)", 
               "Total Emissions (metric tonnes CO2e)",
               "WellBeing scores"),

`Sample mean` = c(energy_t_test$estimate,
                    dens_t_test$estimate,
                    emission_t_test$estimate,
                    wellbeing_t_test$estimate* 1000),

`Population mean (95% confidence) - low` = 
    c(energy_t_test$conf.int[1],
      dens_t_test$conf.int[1],
      emission_t_test$conf.int[1],
      wellbeing_t_test$conf.int[1]* 1000),
`Population mean (95% confidence) - high` =
    c(energy_t_test$conf.int[2],
      dens_t_test$conf.int[2],
      emission_t_test$conf.int[2],
      wellbeing_t_test$conf.int[2] * 1000),
`Median` = c(energy_quartiles[3],
             dens_quartiles[3],
             emission_quartiles[3],
             wellbeing_quartiles[3] * 1000),
`Interquartile range` = c(energy_quartiles[4] - energy_quartiles[2],
                            dens_quartiles[4] - dens_quartiles[2],
                            emission_quartiles[4] - emission_quartiles[2],
                            wellbeing_quartiles[4] - wellbeing_quartiles[2]* 1000),
`Standard deviation` = c(energy_dev,
                          dens_dev,
                          emission_dev,
                          wellbeing_dev* 1000))

kable(cont_summary, digits = 0)
```

```{r display histograms, warning = FALSE, message = FALSE, echo = TRUE, results = FALSE}

pretty_energy_hist <- energy_hist +
  theme_bw() +
  scale_x_continuous(name = "Percentage of Renewable energy") +
  scale_y_continuous(name = "Number of cities") +
  theme(axis.text.x = element_text(angle = 90))

pretty_dens_hist <- dens_hist +
  theme_bw() + 
  scale_x_continuous(name = "Population density\n(residents per square kilometer)",
                     trans = "log",
                     breaks = c(0.1, 1, 10, 100, 1000, 10000),
                     labels = c("0.1", "1", "10", "100", "1000", "10,000")) +
  scale_y_continuous(name = "Number of cities") +
  theme(axis.text.x = element_text(angle = 90))

pretty_emission_hist = emission_hist +
  theme_bw() +
  scale_x_continuous(name = "Total Emissions",
                     breaks = seq(0, 0.3, by=.05),
                     labels = seq(0, 300, by = 50)) +
  scale_y_continuous(name = "Number of cities") +
  theme(axis.text.x = element_text(angle = 90)) 
  
pretty_wellbeing_hist = wellbeing_hist +
  theme_bw() +
  scale_x_continuous(name = "WellBeing scores",
                     breaks = seq(0, 0.3, by=.05),
                     labels = seq(0, 300, by = 50)) +
  scale_y_continuous(name = "Number of cities") +
  theme(axis.text.x = element_text(angle = 90))

grid.arrange(pretty_energy_hist, pretty_dens_hist, pretty_emission_hist, pretty_wellbeing_hist,
             ncol = 2, nrow = 2)
```
# Sample Proportion for each Category

## Our categorical variables are:

* Vulnerability Assessment
* C40 Status

```{r}
pct_C40 <- t.test(main$C40_Status == "C40")
pct_nonC40 <- t.test(main$C40_Status == "No")
```

```{r}
cat_summary_C40 <- tibble(`C40 Cities` = 
                              c("C40",
                                "Non C40"),
                            `Sample proportion` = 
                              c(pct_C40$estimate * 100,
                                pct_nonC40$estimate *100),
                            `95-percent confidence - low` = 
                              c(pct_C40$conf.int[1] * 100,
                                pct_nonC40$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_C40$conf.int[2] * 100,
                                pct_nonC40$conf.int[2] * 100))

kable(cat_summary_C40, digits = 0)
```
# The 95-percent confidence interval for the population proportion in each category

```{r}
pct_VAdone <- t.test(main$Vadone == "Yes")
pct_VAinprogress <- t.test(main$Vadone == "In progress")
pct_VAno <-  t.test(main$Vadone == "No")
```

```{r}

cat_summary_VAdone <- tibble(`Conducted Vulnerability Assessment?` = 
                              c("Yes",
                                "In Progress",
                                "No"),
                            `Sample proportion` = 
                              c(pct_VAdone$estimate * 100,
                                pct_VAinprogress$estimate * 100,
                                pct_VAno$estimate * 100),
                            `95-percent confidence - low` = 
                              c(pct_VAdone$estimate * 100,
                                pct_VAinprogress$estimate * 100,
                                pct_VAno$estimate * 100),
                            `95-percent confidence - high` = 
                              c(pct_VAdone$conf.int[2] * 100,
                                pct_VAinprogress$conf.int[2] * 100,
                                pct_VAno$conf.int[2] * 100))

kable(cat_summary_VAdone, digits = 0)
```

# Bivariate analysis
## Correlations / regression on continuous variables
Well-being is our key variable. The components of the well-being index are

* Annual hours of sunshine hours
* Access to green space
* Crime
* Traffic congstion
* Happiness
* Quality of healthcare
* Work-life Balance
* Governance 

We recognize the limitations of this index and remain critical of its representation of the quality of life of citizens in the cities examined. Nevertheless, it is an interesting study and we are curious if the continuous and categorical variables we've identified can give the index more context as we establish its relationships to these variables.

In this exercise, we would like to investigate the correlations between 

* wellbeing and emissions
* wellbeing and population density
* wellbeing and energy-mix
* population and emissions 
* emissions and energy-mix

## Correlation wellbeing and emissions

The p-value is more than 0.05 at 0.3891, which means the relationship is insignificant. This is also clear from the fact that the entire 95-percent confidence interval is both positive and negative, which suggests that the relationship is neither inverse or direct.
```{r}
cor.test(~ WellBeing + TotalEmissions, data = main)
```

```{r}
wb_emissions_model <- lm(WellBeing ~ TotalEmissions, data = main)

summary(wb_emissions_model)
```

```{r}
ggplot(main) +
  geom_point(aes(x = TotalEmissions, 
                 y = WellBeing,
                 size = TotalEmissions),
             alpha = 0.25) +
  theme_bw()
```

## Correlation wellbeing and population density

The p-value is more than 0.05 at 0.3906, which means the relationship is insignificant. This is also clear from the fact that the entire 95-percent confidence interval is both positive and negative, which suggests that the relationship is neither inverse or direct.
```{r}
cor.test(~ WellBeing + PopulationDensity, data = main)
```

```{r}
wb_popdens_model <- lm(WellBeing ~ PopulationDensity, data = main)

summary(wb_popdens_model)
```

## Correlation/Regression wellbeing and energy mix

The p-value is less than 0.05 at 0.000009851, which means the relationship is significant. This is also clear from the fact that the entire 95-percent confidence interval is positive, which suggests that the relationship is direct.

This result is interesting and it could be suggesting that cleaner cities provide better conditions for a higher wellbeing.

```{r}
cor.test(~ WellBeing + PercentageRenewable, data = main)
```

```{r}
wb_energy_model <- lm(WellBeing ~ PercentageRenewable, data = data)

summary(wb_energy_model)
```


```{r}
ggplot(main) +
  geom_point(aes(x = PercentageRenewable, 
                 y = WellBeing),
             alpha = 0.25) +
  theme_bw()
```
## Correlation population density and emissions

The p-value is more than 0.05 at 0.4849, which means the relationship is insignificant. This is also clear from the fact that the entire 95-percent confidence interval is both positive and negative, which suggests that the relationship is neither inverse or direct.
```{r}
cor.test(~ PopulationDensity + TotalEmissions, data = main)
```

```{r}
popdens_te_model <- lm(PopulationDensity ~ TotalEmissions, data = main)

summary(popdens_te_model)
```

## Correlation emissions and energy mix

The p-value is less than 0.05 at 0.0105, which means the relationship is significant. This is also clear from the fact that the entire 95-percent confidence interval is negative (it does not include zero), which suggests that the relationship is inverse.

```{r}
cor.test(~ TotalEmissions + PercentageRenewable, data = data)
```

```{r}
te_energy_model <- lm(TotalEmissions ~ PercentageRenewable, data = main)

summary(te_energy_model)
```


# Difference in means / regression on categorical variables

## Difference in means on wellbeing and c40 status

The p-value is more than 0.05 at 0.2037, which means the relationship is insignificant. This is also clear from the fact that the entire 95-percent confidence interval is both positive and negative, which suggests that the relationship is neither inverse or direct.

```{r}
t.test(WellBeing ~ C40_Status, main)
```


```{r}
wb_c40_model <- lm(WellBeing ~ C40_Status, data = main)

summary(wb_c40_model)

```


## Difference in means on wellbeing and status of vulnerability assessment (done or not)

#```{r}
t.test(WellBeing ~ Vadone, main)
```
The p-value is less than 0.05 at 0.03, which means the relationship is significant. This is also clear from the fact that the entire 95-percent confidence interval is negative (it does not include zero).

It appears that there cities that have conducted vulnerability assessments are also cities whose citizens have a higher well being score. This could be due to the cities/governments which conduct vulnerability assessments are also more proactive in ensuring the wellbeing of their citizens.

```{r}
wb_va_model <- lm(WellBeing ~ Vadone, data = main)

summary(wb_va_model)

```

